Inferring Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-07-01 doc/paper-lcr.tex
finished introduction draft

\title[Line of Sight Mass Reconstruction]
{Inferring Lensing Mass in the Universe from Photometric Catalog Data}

@@

What is the problem - reconstructing kappa. Why is this problem relevant? Time
delay distances. Magnification.

Previous work on environments: Auger, Momcheva,
others. Hilbert et al.'s raytracing plus Fassnacht \& Koopmans density, Suyu et
al's kappa pdf.

Why these approaches may be insufficient - information provided in
future surveys.

Questions asked in this paper:
How much of the total external convergence in a time delay lens system comes from visible galaxies? How does this change as a function of magnitude cut?
Where are the most important galaxies? Do the massive galaxies dominate, or is it the the galaxies nearest the optical axis, or some combination of these? How many galaxies account for 50%, and 95% of the convergence on a typical line of sight?
Can the Hilbert (ray-traced) convergence be recovered by halo model lightcone reconstruction? With what scatter and bias? Which recipes are the dominant sources of these uncertainties?

This paper is organized as follows. \Sref{sec:}.

@@

\section{Tests with known halo mass and redshift}
\label{sec:}

@@

\section{Tests on mock catalogs}
\label{sec:}

@@

\section{Discussion}
\label{sec:}

Why reconstruction works or doesn't work
Major sources of uncertainty
Ideas for improving the procedure

\section{}
\label{sec:}

@@

\section{}
\label{sec:}

@@

\section{Summary}
\label{sec:summary}

@@

>>>>

\title[Line of Sight Mass Reconstruction]
{Inferring Lensing Mass in the Universe \\
from Photometric Catalog Data}

@@

Averaged over large areas of sky, the weak gravitational lensing effect can be
used to constrain the statistical properties of the matter density in the
Universe: every distant object we observe has had its shape distorted, and size
and total brightness magnified (or demagnified) by a compound lens constructed
from all the mass distributed between us and it.

This fact makes gravitational lensing a potentially important source of
systematic error for any estimate of luminosity or distance, an issue addressed
for \eg type Ia supernovae by \citet[][]{Holz+Wald,Linder+Holz} and for high
redshift galaxies by \citet{Loeb} among others. In time delay lens cosmography
the effect can be particularly important. \citet{Suyu2010} found that  the
lensing effect of mass structures along the (unusually over-dense)  line of
sight to the quadruply-imaged radio source B1608$+$656, the so-called ``external
convergence,'' was the dominant source of uncertainty in the inferred
cosmological parameters.

How can we correct for the line of sight lensing error, and make accurate
measurements through the Universe? As \citeauthor{LH} noted, when inferring
global quantities (such as the cosmological parameters), averaging the results
from a large number of independent objects will tend to reduce the error, as the
ensemble-averaged convergence approaches zero for an unbiased sample
\citep{xxx}. If the sample is biased, however, some residual systematic error
will remain while the statistical uncertainty decreases -- and this will be the
case for any sample selected by the brightness of the lensed object, or any
other quantity that depends on $\xkappa$. The latter class includes time delay
lenses, which are detected based on the separation of their images, which
increases with $\kappa$. Large samples of lenses are expected to be discovered
in the next decade in ground-based optical imaging surveys
\citep{Oguri+Marshll2010} -- but these samples will be image separation-limited,
due to the surveys' relatively poor ($\sim 1$~arcsec) image quality. Some of the
current sample of known time delay lenses were detected in surveys like this: we
might expect them to be somewhat biased towards high values of $\xkappa$.

There are two broad approaches one could take at this point. First, one could
seek to characterise the lensed objects' selection function, and use this
understanding to derive a suitable correction to cosmological parameters
inferred assuming no $\xkappa$ at all. To do this, one needs information about
the unlensed population of objects, and the population of lenses, and how the
objects were detected -- or to infer these hierarchically during the analysis.
Alternatively, one could simply accept all lenses provided, regardless of their
selection, and try to use any additional information that is available to
estimate the external convergence in each system on a case by case basis, and
hence optimize their joint analysis to produce a better-constrained result. In
this work we follow the second approach, and use measurements of the positions,
magnitude and colours of galaxies close to time delay lens lines of sight to try
to infer the external convergence at the lens position.

The first attempts at reconstructing the line of sight mass distribution in
strong lens fields were \ldots. \flag{More needed here. Previous work on
environments: Keeton, Zabludoff, Auger, Momcheva, Wong, others. Spectroscopic
surveys, modelling mass along line of sight.}

These observational investigations provided useful insight into the problem. 
Coming from the opposite direction, \citet{Hilbert++2009} performed ray-tracing
experiments through the Millenium Simulation, generating a predicted $\xkappa$
at every position in a simulated sky. The first step towards using this
excellent calibration resource was taken by \citet{Suyu++2010}, who selected
Millenium lines of sight that a) were found to contain a strong lens and b) had
comparable apparent galaxy overdensity in a 45 arcsec radius aperture down to $i
< 24$. This overdensity could be compared to the overdensity of the B1608$+$656
line of sight \citep{Fass+Koop2010}, and hence an estimate of the probability
density function (PDF) $\Pr(\xkappa)$ could be made. This PDF is the key
quantity for any ensemble analysis, since by marginalising over it we can
correctly propagate the uncertainties due to weak lensing by line of sight mass
structure.

In this paper we use the Millenium Simulation mock galaxy catalogs and
ray-traced lensing effects to test the accuracy of line of sight mass
reconstructions. Probabilistically ssigning mass and redshift to every observed
galaxy in the field around a time delay lens, we generate Monte Carlo sample
line of sight mass distributions, and estimate the external convergence due to
each one. This gives us a $\Pr(\xkappa)$ that can be propagated into further
joint inferences. We focus on time delay lens cosmography, and quantify our
results in terms of the uncertainty on the mean time delay distance to a test
sample of $N$ lenses (where for clarity each lens has the same deflector and
source redshift, and the same lens model and time delay uncertainty).

We aim to answer the following questions:
\begin{itemize}
\item Faint galaxies and other dark structures will not appear in a photometric
object catalog, but they will contribute convergence at some level. How much of
the total external convergence in a time delay lens system comes from visible
galaxies? How does this change as a function of magnitude cut? 
\item Can the true (ray-traced) convergence be recovered by halo model
reconstruction? What scatter and residual bias are induced in the ensemble 
distance estimate? Which aspects of the model dominate these uncertainties? 
\item When faced with a newly-detected lens, surrounded by galaxies on the sky,
we have some choices to make when planning follow-up observations.
Which are the most important galaxies, with regard to the external convergence
produced? Can nearby groups and clusters be straightforwardly accounted for? If
lenses with such massive structures nearby are discarded, what impact does that
have on the distance accuracy?
\end{itemize}

This paper is organized as follows. In~\Sref{sec:theory} we review the relevant
gravitational lensing theory, before introducing our simple reconstruction 
model in \Sref{sec:model}.  We then test this model in two phases: first, in
\Sref{sec:knownMh+z}, with known redshift and halo mass for every galaxy in a
lens field, in order to quantify the irreducible uncertainty due to unseen mass,
and second, in \Sref{sec:obsMstar+z}, with realistic observational uncertainties
on the field galaxies' stellar masses and redshifts. We then propagate all these
uncertainties through to the time delay distance in a toy sample of 100 lenses,
in \Sref{sec:distances}. We discuss our results in \Sref{sec:discuss} before
concluding in \Sref{sec:conclude}.

@@

\section{Tests with known halo mass and redshift}
\label{sec:knownMh+z}

@@

\section{Tests on mock catalogs}
\label{sec:obsMstar+z}

@@

Also: compare with simple averaging. This will fail if lenses are selected to
live on over-dense lines of sight. Stefan's plot in Suyu et al shows effect of
selection - what is the resulting bias? Few percent? This is the target to beat,
need to find it out.

Make mock sample of lenses, all wit B1608 zd and zs, but with different kappas.
Draw kappas from Stefan's *lens* distribution. Can we correct for the selection
effect?

@@

\section{Discussion}
\label{sec:discuss}

Why reconstruction works or doesn't work.

Major sources of uncertainty.

Ideas for improving the procedure

@@

\section{Conclusions}
\label{sec:conclude}

To remind you, this is what we did.

@@

% \item Faint galaxies and other dark structures will not appear in a photometric
% object catalog, but they will contribute convergence at some level. How much of
% the total external convergence in a time delay lens system comes from visible
% galaxies? How does this change as a function of magnitude cut?

% \item Can the true (ray-traced) convergence be recovered by halo model
% reconstruction, and with what scatter and residual bias? Which aspects of the 
% model dominate these uncertainties?

% \item When faced with a newly-detected lens, surrounded by galaxies on the sky,
% we have some choices to make when planning follow-up observations.
% Which are the most important galaxies, with regard to the external convergence
% produced? Can nearby groups and clusters be straightforwardly accounted for? If
% lenses with such massive structures nearby are discarded, what impact does that
% have on the distance accuracy?
